---
title: "STAT184 Final Project Jake-Part"
author: "Jake Silva"
date: "2024-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background on the Dataset

The data set we decided to use for this analysis is called FIFA 24 Player Stats Data set. We found this set through Kaggle.com. This entire data set had a total of 6 CSVs including Male Coaches, Male Teams, Male Players, Female Coaches, Female Teams, Female Players. All our data in this report comes from the Male Players CSV and selected all the data for the year 2024. Inside this set, there was a 109 attributes ranging from the players date of birth, nationality, position, ranking statistics, and more. As goalkeepers in-game statistics are weighted differently, the scope of this report is just field players. In FIFA, the game displays players using 6 main statistics which include a players rating for Pace, Shooting, Passing, Dribbling, Defense, and Physicality. In this project, the main attributes we focused on where Player Name, Market Value, Overall Rating, Position, and the 6 main statistics mentioned above. Moreover, a player defines a case through this analysis.


## Data Visualization #1
We want to answer the question of how does a players overall rating compare to their market value. Intuitively, we suspect a positive correlation between these variables; however, we want to see just how strong that correlation might be. Additionally, what rating to the price of the players start to vary. A scatterplot is the best option to answer all of these questions.

```{r Visualization 1, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

player_stats <- read_csv("/Users/jakesilva/Downloads/male_players.csv")

#Clean the data of Null Values
player_stats_clean <- player_stats %>%
  remove_empty(c("rows", "cols")) %>%  
  drop_na(club_position, player_positions, overall, value_eur)  

#Filtering to exclude goalkeepers and 2024
field_players <- player_stats_clean %>%
  filter(club_position != "GK", player_positions != "GK", fifa_version == 24.0)

#Plot
ggplot(player_stats, aes(x = overall, y = value_eur)) +
  geom_point(alpha = 0.3, color = "blue", size = 1.5) +  
  labs(x = "Overall Rating", y = "Market Value (EUR)", title = "Overall Rating vs Market Value") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 19, face = "bold", hjust = 0.5),  
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 15),  
    axis.text.y = element_text(size = 15)   
  ) +
  scale_y_log10(labels = scales::comma) + 
  geom_smooth(method = "lm", se = FALSE, color = "red")
```

As we can see, there is a strong positive correlation between a player's overall rating and their market value. This is demonstrated by the red regression line with a positive constant slope. The areas of darker blue have more players in that area, and the lighter the blue dot the less players. Additionally, the players with lower overall rating (50-65) tend to vary more in price than the higher overall ratings(80-90).


## Data Visualization #2
To answer our second research question we want to perform a statistical analysis on Messi vs Ronaldo. If you've heard of the sport of soccer, chances are you've heard these two names as well. The media portrays Messi and Ronaldo as the best players of all time. This creates division among the soccer community as there is a common debate of who is actually better than the other. For the scope of this project, we want to observe who is the better player in game. Through a side-by-side histogram, we will analyze their 6 main statistics, as well as their overall ranking.

```{r Data Visualization 2, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
player_stats <- read_csv("/Users/jakesilva/Downloads/male_players.csv")

#Clean the data of Null Values
player_stats_clean <- player_stats %>%
  remove_empty(c("rows", "cols")) %>%  
  drop_na(club_position, player_positions, overall, value_eur)

#Get Messi and Ronaldo's Player ID
comparison_data <- player_stats_clean %>%
  filter(fifa_version == 24.0, player_id %in% c(158023, 20801))

#Get Attributes for comparisons
comparison_data_long <- comparison_data %>%
  select(player_id, long_name, overall, defending, pace, shooting, passing, dribbling, physic) %>%
  pivot_longer(cols = c(overall, defending, pace, shooting, passing, dribbling, physic), 
               names_to = "attribute", values_to = "rating") %>%
#Change the attribute labels
  mutate(attribute = factor(attribute,
                            levels = c("pace", "shooting", "passing", "dribbling", "defending", "physic", "overall"),
                            labels = c("Speed", "Shooting", "Passing", "Dribbling", "Defense", "Physical", "Overall Rating")))

player_colors <- c("Lionel Andrés Messi Cuccittini" = "blue", "Cristiano Ronaldo dos Santos Aveiro" = "red")

#Plot
ggplot(comparison_data_long, aes(x = attribute, y = rating, fill = long_name)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.7) +
  scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  scale_fill_manual(values = player_colors, 
                    name = "Player",  
                    labels = c("Cristiano Ronaldo", "Lionel Messi")) +
  labs(title = "Cristiano Ronaldo vs Lionel Messi",
       x = "", y = "Rating") +
  theme_minimal() + 
  #Legend
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ attribute, scales = "free_x", ncol = 7) 
```

This histogram allows to easily visualize how Messi and Ronaldo compare side-by-side. Messi takes the edge on Ronaldo in Dribbling, Pace, and Passing. While Ronaldo is statistically better at Defending, Shooting, and Physicality. As for the overall answer we are looking for... Messi is a higher overall rated player in FIFA.

